---
title: "BCI 50-hectare plot scaling"
author: "Quentin D. Read"
date: "April 26, 2017"
output: 
  pdf_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This document shows how the scaling relationships for the BCI 50-hectare plot are calculated, and outputs all the figures. John's proposed figures 1 and 2 are schematics, and figures 3, 4, and supplemental are with data. For each relationship, I plot all trees together, then I plot gap, intermediate, and shade groups separately. I also do the scaling by diameter and mass and plot each of them. We may not use the mass scaling figures but I have them in the document in the interest of completeness. Here, I use the two most recent BCI surveys (survey 6 and survey 7). This can be changed if desired. Also, the light availability relationships are based on a crowding index and not on light availability directly, because the latter is much more difficult to estimate. Those results are pending.

\newpage
# Functions needed to calculate scalings

## Fit power law

```{r}
powerlawfit <- function(dat) {
  library(poweRlaw)
  pl_dat <- conpl$new(dat)
  lognorm_dat <- conlnorm$new(dat)
  xmin_pl <- pl_dat$getXmin()
  xmin_lognorm <- lognorm_dat$getXmin()
  pars_pl <- estimate_pars(pl_dat)
  pars_lognorm <- estimate_pars(lognorm_dat)
  pl_dat$setPars(pars_pl)
  lognorm_dat$setPars(pars_lognorm)
  plotdat <- plot(pl_dat)
  plfit_dat <- lines(pl_dat)
  lognormfit_dat <- lines(lognorm_dat)
  pl_pdf <- dist_pdf(m = pl_dat, q = plfit_dat$x, log = FALSE)
  lognorm_pdf <- dist_pdf(m = lognorm_dat, q = lognormfit_dat$x, log = FALSE)
  
  # bootstrap confidence interval of Pareto fit
  # discard 500 burnin iterations
  n_boot <- 1499
  n_burn <- 500
  pl_boot <- bootstrap(m = pl_dat, xmins = pl_dat$getXmin(), no_of_sims = n_boot, threads = 3)
  boot_ci <- quantile(pl_boot$bootstraps$pars[-(1:n_burn)], probs = c(0.025, 0.975))
  
  return(list(plotdat = plotdat, 
              plfit = plfit_dat, 
              lognormfit = lognormfit_dat, 
              plpdf = data.frame(x = plfit_dat$x, y = pl_pdf),
              lognormpdf = data.frame(x = lognormfit_dat$x, y = lognorm_pdf),
              xmin = xmin_pl, 
              alpha = pars_pl$pars,
              xmin_lognorm = xmin_lognorm,
              pars_lognorm = pars_lognorm$pars,
              boot_ci = as.numeric(boot_ci)
              ))
}
```

\newpage
## Log-bin algorithm

```{r}
logbin <- function(x, y = NULL, n) {
  logx <- log10(x)                                           # log transform x value (biomass)
  bin_edges <- seq(min(logx), max(logx), length.out = n + 1) # get edges of bins
  logxbin <- rep(NA, length(logx))                           # create data structure to assign trees to bins
  b <- bin_edges                                             # add a little to the biggest bin temporarily
  b[length(b)] <- b[length(b)] + 1                           # (so that the biggest single tree is put in a bin)
  for (i in 1:length(logx)) {
    logxbin[i] <- sum(logx[i] >= b)                          # assign each tree to a bin
  }
  bin_midpoints <- numeric(n)
  for (i in 1:n) {
    bin_midpoints[i] <- mean(10^(bin_edges[i:(i+1)]))        # backtransform bin edges to linear, and get midpoints
  }
  bin_widths <- diff(10^bin_edges)                           # get linear width of each bin
  bin_factor <- factor(logxbin, levels=1:n)                  # convert bin to factor (required to deal with zeroes if present)
  bin_counts <- table(bin_factor)                            # find number of trees in each bin
  if (!is.null(y)) {
    rawy <- tapply(y, bin_factor, sum)                       # sum y value (production) in each bin
    rawy[is.na(rawy)] <- 0                                   # add zeroes back in if present
    bin_values <- as.numeric(rawy/bin_widths)                # divide production by width for each bin 
  }
  else {
    bin_values <- as.numeric(bin_counts/bin_widths)          # 1-dimensional case.
  }
  
  return(data.frame(bin_midpoint = bin_midpoints,            # return result!
                    bin_value = bin_values,                  # also add bin min and max for bar plot purposes
                    bin_count = as.numeric(bin_counts),
                    bin_min = 10^bin_edges[1:n],
                    bin_max = 10^bin_edges[2:(n+1)]))
  
}
```

\newpage
## Plot log bins as histogram

```{r}
plotlogbin <- function(dat, xl, yl, plottitle, reg = FALSE) {
  
  y_min <- 10^floor(log10(min(dat$bin_value, na.rm = TRUE)))
  y_max <- max(dat$bin_value, na.rm = TRUE) * 1.1
  
  p <- ggplot(dat, aes(xmin=bin_min, xmax=bin_max, ymin=0, ymax=bin_value)) + 
    geom_rect(alpha = 0.5) +
    scale_x_log10(name = xl, expand = c(0,0),
                  breaks = scales::trans_breaks("log10", function(x) 10^x),
                  labels = scales::trans_format("log10", scales::math_format(10^.x))) +
    scale_y_log10(name = yl, expand = c(0,0), limits = c(y_min, y_max),
                  breaks = scales::trans_breaks("log10", function(x) 10^x),
                  labels = scales::trans_format("log10", scales::math_format(10^.x))) +
    panel_border(colour = 'black') + 
    ggtitle(plottitle)
  if (reg) {
    p <- p +
      stat_smooth(method = 'lm', se = FALSE, color = 'black', 
                  aes(x = bin_midpoint, y = bin_value)) +
      geom_text(x = -Inf, y = -Inf, 
                label = paste('Slope:', 
                              round(lm(I(log10(bin_value)) ~ I(log10(bin_midpoint)), data=dat)$coef[2], 2)),
                hjust = 0, vjust = -0.25)
  }
  return(p)
}
```

\newpage
## Plot log bins and the power law fit on one plot

```{r}
plotbinsandfits <- function(pl, bindat, plottitle = 'plot title', xl = 'x label', yl = 'log PDF') {
  
  expr1 <- as.character(as.expression(substitute(
    "Pareto:"~~alpha == a, list(a = round(pl$alpha, 2)))))

  bindat <- transform(bindat, bin_value = bin_value/sum(bin_count))
  bindat <- subset(bindat, bin_value > 0)
  y_min <- 10^floor(log10(min(bindat$bin_value, na.rm = TRUE)))
  y_max <- max(bindat$bin_value, na.rm = TRUE) * 1.1
  
  p <- ggplot(bindat) + 
    scale_x_log10(name = xl,
                  breaks = scales::trans_breaks("log10", function(x) 10^x),
                  labels = scales::trans_format("log10", scales::math_format(10^.x)),
                  expand = c(0,0)) +
    scale_y_log10(name = yl,
                  limits = c(y_min, y_max),
                  breaks = scales::trans_breaks("log10", function(x) 10^x),
                  labels = scales::trans_format("log10", scales::math_format(10^.x)),
                  expand=c(0,0)) +
    geom_rect(aes(xmin = bin_min, xmax = bin_max, ymin = 0, ymax = bin_value), alpha = 0.5) +
    geom_line(data = subset(pl$plpdf, y>0), aes(x,y), color = 'forestgreen', size = 2) +
    geom_text(x = -Inf, y = -Inf, label = expr1, parse = TRUE, hjust = 0, vjust = -1.5) +
    panel_border(colour = 'black') +
    labs(x = xl, y = yl) +
    ggtitle(plottitle)
  return(p)
}
```

\newpage
## Plot adjusted and unadjusted fits on scatterplot

```{r}
twoslopeplot <- function(dat, plottitle = 'plot title', xl = 'x label', yl = 'y label', binvar='diameter') {
  if (binvar == 'mass') dat$yv <- dat$agb else dat$yv <- dat$dbh
  lmsize <- lm(log10(production67) ~ log10(yv), data=dat)
  lmsizecomp <- lm(log10(production67) ~ log10(yv) + log10(comp_idx), data=dat)
  dbh_slope <- lmsizecomp$coefficients[2] # extract slope from full model
  # refit model, setting slope from full model and estimating intercept
  adjusted_lm <- lm(log10(production67) ~ 1 + offset(dbh_slope * log10(yv)), data=dat) 
  adjusted_intercept <- adjusted_lm$coefficients[1]
  
  ggplot(dat, aes(x = yv, y = production67)) +
    geom_point(alpha = 0.6) +
    geom_abline(slope = lmsize$coefficients[2], intercept = lmsize$coefficients[1], color = 'indianred', size = 2) +
    geom_abline(slope = dbh_slope, intercept = adjusted_intercept, color = 'steelblue1', size = 2) +
    scale_x_log10(name = xl,
                  breaks = scales::trans_breaks("log10", function(x) 10^x),
                  labels = scales::trans_format("log10", scales::math_format(10^.x))) + 
    scale_y_log10(name = yl,
                  breaks = scales::trans_breaks("log10", function(x) 10^x),
                  labels = scales::trans_format("log10", scales::math_format(10^.x))) +
    ggtitle(plottitle) +
    panel_border(colour = 'black')
}
```


\newpage
# Preparing data

The competition index was already calculated elsewhere. Ask me if you want the code for it.

```{r, message = FALSE, warning = FALSE}
library(dplyr)
library(cowplot)

fp <- 'C:/Users/Q/Dropbox/projects/forestlight/bcidata/'
bcicensusdat <- read.csv(file.path(fp, 'bci_compidx.csv'), stringsAsFactors = FALSE)

# Convert tonnes to kg
bcicensusdat <- mutate(bcicensusdat,
                       agb=agb*1000,
                       production67=production67*1000)

```

Join the BCI census data with categorical and continuous shade tolerance based on Wright's data.

```{r, echo=F, message=F}
# Load Wright's growth data and back calculate the quantiles for shade tolerance
library(XLConnect)
wright <- readWorksheetFromFile(file = 'C:/Users/Q/Google Drive/ForestLight/data/Shade Tolerance/Demographic/Wright et al 2010, growth mortality tradeoffs.xlsx', sheet = 1, startRow = 26) # Get rid of the lines above header.
wright[wright == -99] <- NA # Unknown values were given a value of -99
wright$SPECIES.[109:110] <- c('simplex_var1', 'simplex_var2') # Correct duplicate named subspecies.
wright$Taxon <- with(wright, paste(GENUS., SPECIES.))

wright_df <- with(wright, data.frame(Taxon, mrt = MRT25SAP/100, rgr = RGR95SAP, stringsAsFactors = FALSE))
wright_df <- subset(wright_df, !is.na(mrt))

wright_pca <- with(wright_df, prcomp(data.frame(qlogis(mrt), log10(rgr)), scale=TRUE, center=TRUE)) # 90% of variation on the growth-mortality single axis. Nice.
pca_scores <- wright_pca$x[,1]
pca_groups <- cut(pca_scores, breaks = 3)
pca_groupcodes <- factor(pca_groups, labels = c('G','I','S'))

wright_df <- data.frame(wright_df, pca_scores, tol_wright = as.character(pca_groupcodes), stringsAsFactors = FALSE)

# Correct wright_df entries that are not correct.
wright_df$Taxon[grep('Beilsc',wright_df$Taxon)] <- 'Beilschmiedia pendula'
wright_df$Taxon[grep('Cestrum',wright_df$Taxon)] <- 'Cestrum megalophyllum'
wright_df$Taxon[grep('phyllu arg',wright_df$Taxon)] <- 'Chrysophyllum argenteum'
wright_df$Taxon[grep('Coccol',wright_df$Taxon)] <- 'Coccoloba manzinellensis'
wright_df$Taxon[grep('Tabern',wright_df$Taxon)] <- 'Tabernaemontana arborea'
wright_df$Taxon[grep('var1',wright_df$Taxon)] <- 'Swartzia simplex_var.grandiflora'
wright_df$Taxon[grep('var2',wright_df$Taxon)] <- 'Swartzia simplex_var.ochnacea'
wright_df$Taxon[grep('colorado',wright_df$Taxon)] <- 'Eugenia coloradoensis'
wright_df$Taxon[grep('Croton',wright_df$Taxon)] <- 'Croton billbergianus'
```


```{r}
bci_lookup <- read.delim('C:/Users/Q/Dropbox/projects/forestlight/bcidata/ViewTax.txt', stringsAsFactors = FALSE)

bci_lookup <- bci_lookup %>%
  mutate(Taxon = paste(Genus, SpeciesName)) 

taxmatch <- bci_lookup$Taxon %in% wright_df$Taxon

bcicensusdat <- left_join(bci_lookup, wright_df) %>%
  rename(sp = Mnemonic) %>%
  select(sp, mrt, rgr, pca_scores, tol_wright) %>%
  right_join(bcicensusdat)

```

\newpage
# Scaling relationships based on diameter

## Individual production

These plots are figure 3, left column, in John's mockups.

```{r}
xl1 <- 'Diameter (cm)'
yl1 <- expression(paste('Individual production (kg y'^-1,')', sep=''))

alltreedat <- subset(bcicensusdat, !is.na(dbh) & production67 > 0 & comp_idx > 0)
twoslopeplot(dat = alltreedat, 
             plottitle = 'All species', 
             xl = xl1, 
             yl =  yl1)

shadedat <- subset(bcicensusdat, !is.na(dbh) & production67 > 0 & comp_idx > 0 & tol_wright == 'S')
twoslopeplot(dat = shadedat, 
             plottitle = 'Shade-tolerant species', 
             xl = xl1, 
             yl =  yl1)

intdat <- subset(bcicensusdat, !is.na(dbh) & production67 > 0 & comp_idx > 0 & tol_wright == 'I')
twoslopeplot(dat = intdat, 
             plottitle = 'Intermediate species', 
             xl = xl1, 
             yl =  yl1)

gapdat <- subset(bcicensusdat, !is.na(dbh) & production67 > 0 & comp_idx > 0 & tol_wright == 'G')
twoslopeplot(dat = gapdat, 
             plottitle = 'Gap species', 
             xl = xl1, 
             yl =  yl1)
```


\newpage
## Binned density

These plots are figure 3, right column, in John's mockups. 

```{r, results='hide', fig.keep = 'none', warning = FALSE, message = FALSE}
numbins <- 20 # Can be edited if desired. Only for visualization purposes.

# Pareto fits
bci_dens_fit_all <- powerlawfit(alltreedat$dbh)
bci_dens_fit_shade <- powerlawfit(shadedat$dbh)
bci_dens_fit_inter <- powerlawfit(intdat$dbh)
bci_dens_fit_gap <- powerlawfit(gapdat$dbh)

# Log bins
bci_dens_logbin_all <- with(alltreedat, logbin(x=dbh, y=NULL, n = numbins))
bci_dens_logbin_shade <- with(shadedat, logbin(x=dbh, y=NULL, n = numbins))
bci_dens_logbin_inter <- with(intdat, logbin(x=dbh, y=NULL, n = numbins))
bci_dens_logbin_gap <- with(gapdat, logbin(x=dbh, y=NULL, n = numbins))
```

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.width = 3.25, fig.height = 3.25}
xl2 <- 'Diameter (cm)'
yl2 <- 'Density'

plotbinsandfits(bci_dens_fit_all, bci_dens_logbin_all,
                plottitle = 'All species', xl = xl2, yl = yl2)

plotbinsandfits(bci_dens_fit_shade, bci_dens_logbin_shade,
                plottitle = 'Shade-tolerant species', xl = xl2, yl = yl2)

plotbinsandfits(bci_dens_fit_inter, bci_dens_logbin_inter,
                plottitle = 'Intermediate species', xl = xl2, yl = yl2)

plotbinsandfits(bci_dens_fit_gap, bci_dens_logbin_gap,
                plottitle = 'Gap species', xl = xl2, yl = yl2)
```


\newpage
## Binned production

These plots are figure 4, bottom row, in John's mockups. (The top row will be added once I calculate light availability.)

```{r, results = 'hide', fig.keep = 'none'}
numbins <- 20 # Can be edited if desired. ***NOT JUST FOR LOOKS***

# Log bins
bci_prod_logbin_all <- with(alltreedat, 
                   logbin(x=dbh, y=production67, n=numbins))
bci_prod_logbin_shade <- with(shadedat, 
                         logbin(x=dbh, y=production67, n=numbins))
bci_prod_logbin_inter <- with(intdat, 
                         logbin(x=dbh, y=production67, n=numbins))
bci_prod_logbin_gap <- with(gapdat, 
                       logbin(x=dbh, y=production67, n=numbins))

# Energy-equivalence slopes
bci_prod_lm_all <- lm(log10(bin_value) ~ log10(bin_midpoint), data = bci_prod_logbin_all)
bci_prod_lm_shade <- lm(log10(bin_value) ~ log10(bin_midpoint), data = bci_prod_logbin_shade)
bci_prod_lm_inter <- lm(log10(bin_value) ~ log10(bin_midpoint), data = bci_prod_logbin_inter)
bci_prod_lm_gap <- lm(log10(bin_value) ~ log10(bin_midpoint), data = bci_prod_logbin_gap)

```

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.width = 3.25, fig.height = 3.25}
xl3 <- 'Diameter (cm)'
yl3 <- expression(paste('Total production (kg y'^-1,')', sep=''))

plotlogbin(bci_prod_logbin_all, xl3, yl3, 
           'All species', reg = TRUE)
plotlogbin(bci_prod_logbin_shade, xl3, yl3,  
           'Shade-tolerant species', reg = TRUE)
plotlogbin(bci_prod_logbin_inter, xl3, yl3, 
           'Intermediate species', reg = TRUE)
plotlogbin(bci_prod_logbin_gap, xl3, yl3, 
           'Gap species', reg = TRUE)
```

\newpage
## Slope parameters, with confidence intervals

Below are figures with the slopes and confidence intervals, as well as a table.

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.width = 5, fig.height = 3.25}

lmsize_all <- lm(log10(production67) ~ log10(dbh), data=alltreedat)
lmsizecomp_all <- lm(log10(production67) ~ log10(dbh) + log10(comp_idx), data=alltreedat)
lmsize_shade <- lm(log10(production67) ~ log10(dbh), data=shadedat)
lmsizecomp_shade <- lm(log10(production67) ~ log10(dbh) + log10(comp_idx), data=shadedat)
lmsize_int <- lm(log10(production67) ~ log10(dbh), data=intdat)
lmsizecomp_int <- lm(log10(production67) ~ log10(dbh) + log10(comp_idx), data=intdat)
lmsize_gap <- lm(log10(production67) ~ log10(dbh), data=gapdat)
lmsizecomp_gap <- lm(log10(production67) ~ log10(dbh) + log10(comp_idx), data=gapdat)

slopedat <- data.frame(withcompidx = rep(c('comp idx','no comp idx'), each=4),
                       guild = c('all','shade','intermediate','gap'),
                       slope = c(lmsizecomp_all$coef[2], lmsizecomp_shade$coef[2], lmsizecomp_int$coef[2], lmsizecomp_gap$coef[2],
                                 lmsize_all$coef[2], lmsize_shade$coef[2], lmsize_int$coef[2], lmsize_gap$coef[2]),
                       cimin = c(confint(lmsizecomp_all)[2,1], confint(lmsizecomp_shade)[2,1], confint(lmsizecomp_int)[2,1], confint(lmsizecomp_gap)[2,1],
                                 confint(lmsize_all)[2,1], confint(lmsize_shade)[2,1], confint(lmsize_int)[2,1], confint(lmsize_gap)[2,1]),
                       cimax = c(confint(lmsizecomp_all)[2,2], confint(lmsizecomp_shade)[2,2], confint(lmsizecomp_int)[2,2], confint(lmsizecomp_gap)[2,2],
                                 confint(lmsize_all)[2,2], confint(lmsize_shade)[2,2], confint(lmsize_int)[2,2], confint(lmsize_gap)[2,2]))

ggplot(slopedat, aes(x = guild, y = slope, ymin = cimin, ymax = cimax)) +
  facet_grid(. ~ withcompidx) +
  geom_pointrange() +
  panel_border(colour = 'black') +
  ggtitle('Individual production')
```

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.width = 3.25, fig.height = 3.25}
# Density
slopedat <- data.frame(guild = c('all','shade','intermediate','gap'),
                       slope = c(bci_dens_fit_all$alpha, bci_dens_fit_shade$alpha, bci_dens_fit_inter$alpha, bci_dens_fit_gap$alpha),
                       cimin = c(bci_dens_fit_all$boot_ci[1], bci_dens_fit_shade$boot_ci[1], bci_dens_fit_inter$boot_ci[1], bci_dens_fit_gap$boot_ci[1]),
                       cimax = c(bci_dens_fit_all$boot_ci[2], bci_dens_fit_shade$boot_ci[2], bci_dens_fit_inter$boot_ci[2], bci_dens_fit_gap$boot_ci[2]))

ggplot(slopedat, aes(x = guild, y = slope, ymin = cimin, ymax = cimax)) +
  geom_pointrange() +
  panel_border(colour = 'black') +
  ggtitle('Density by diameter')

# Binned production
slopedat <- data.frame(guild = c('all','shade','intermediate','gap'),
                       slope = c(bci_prod_lm_all$coef[2], bci_prod_lm_shade$coef[2], bci_prod_lm_inter$coef[2], bci_prod_lm_gap$coef[2]),
                       cimin = c(confint(bci_prod_lm_all)[2,1], confint(bci_prod_lm_shade)[2,1], confint(bci_prod_lm_inter)[2,1], confint(bci_prod_lm_gap)[2,1]),
                       cimax = c(confint(bci_prod_lm_all)[2,2], confint(bci_prod_lm_shade)[2,2], confint(bci_prod_lm_inter)[2,2], confint(bci_prod_lm_gap)[2,2]))

ggplot(slopedat, aes(x = guild, y = slope, ymin = cimin, ymax = cimax)) +
  geom_pointrange() +
  panel_border(colour = 'black') +
  ggtitle('Production binned by diameter')

```

### Table of slopes

Guild            | Density | Indiv. production | Density + indiv. production | Binned production | Indiv. production with comp. index
---------------- | ------- | ----------------- | --------------------------- | ----------------- | ----------------------------------
All trees        | -1.88   | 1.93              | 0.05                        | -0.38             | 1.38
Shade-tolerant   | -1.96   | 1.87              | -0.09                       | -0.42             | 1.51
Intermediate     | -1.76   | 1.95              | 0.19                        | 0.18              | 1.25
Gap/pioneer      | -1.99   | 2.05              | 0.06                        | 0.08              | 1.33

### Slope sensitivities to addition of competition index

Guild            | Delta
---------------- | -----
All trees        | 0.55
Shade-tolerant   | 0.36
Intermediate     | 0.70
Gap/pioneer      | 0.72

This supports the hypothesis that sensitivity will increase as shade tolerance decreases.


\newpage
## Shade tolerance as continuous variable

The other analyses above were done separately for each shade tolerance group. An alternative is to use the growth-mortality tradeoff axis (calculated with a PCA) as a continuous variable. We can calculate species means to see whether the species mean of sensitivity to light is related to the species mean position on the axis.

```{r}
getslope <- function(x) {
  xlm <- lm(log10(production67) ~ log10(dbh) + log10(comp_idx), 
            data=x, subset = !is.na(dbh) & production67 > 0 & comp_idx > 0)
  data.frame(n_indiv = with(x, sum(!is.na(dbh) & production67 > 0 & comp_idx > 0)),
             comp_sensitivity = as.numeric(xlm$coefficients[3]))
}


compslopes <- bcicensusdat %>% 
  filter(!is.na(pca_scores)) %>%
  group_by(sp, pca_scores) %>%
  do(getslope(.))
```

Fit a linear model, weighted by the number of individuals per species. The pca score explains a significant proportion of the variation, in the way expected. A higher pca score is associated with shade species, which have higher (closer to zero) sensitivity of production to competition index.

```{r}
sens_lm <- lm(comp_sensitivity ~ pca_scores, data=compslopes, weights = n_indiv)

summary(sens_lm)
```

In the visualization, the size of each point corresponds to the number of individuals per species.

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.width = 3.25, fig.height = 3.25}
ggplot(compslopes, aes(x = pca_scores, y = comp_sensitivity)) +
  geom_hline(yintercept = 0, lty = 3, color = 'gray50', size = 1.5) +
  geom_point(aes(size = log10(n_indiv)), pch = 1) +
  geom_abline(slope=sens_lm$coef[2], intercept = sens_lm$coef[1], color = 'blue', size = 2) +
  scale_size_continuous(name = 'N', breaks=c(2,3,4), labels=10^c(2,3,4)) +
  labs(x = 'Shade tolerance', y = 'Production sensitivity')
```

The same relationship can be tested for density, with a Pareto slope for each species. The question tested here is whether the Pareto slope is steeper as shade tolerance increases (because there are more small saplings of shade-tolerant species)?

```{r}
getparetoslope <- function(x) {
  library(poweRlaw)
  dat <- subset(x, !is.na(dbh) & production67 > 0 & comp_idx > 0)
  pl_dat <- conpl$new(dat$dbh)
  pars_pl <- estimate_pars(pl_dat)
  data.frame(n_indiv = nrow(dat),
             alpha = pars_pl$pars)
}


densslopes <- bcicensusdat %>% 
  filter(!is.na(pca_scores)) %>%
  group_by(sp, pca_scores) %>%
  do(getparetoslope(.))
```

The shade tolerance score does not have any relationship with the slope of the density scaling.

```{r}
densslopes_lm <- lm(alpha ~ pca_scores, data=densslopes, weights = n_indiv)

summary(densslopes_lm)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.width = 3.25, fig.height = 3.25}
ggplot(densslopes, aes(x = pca_scores, y = alpha)) +
  geom_point(aes(size = log10(n_indiv)), pch = 1) +
  scale_size_continuous(name = 'N', breaks=c(2,3,4), labels=10^c(2,3,4)) +
  labs(x = 'Shade tolerance', y = 'Density slope')
```

\newpage
# Scaling relationships based on mass

## Individual production

```{r}
xl1 <- 'Mass (kg)'
yl1 <- expression(paste('Individual production (kg y'^-1,')', sep=''))

alltreedat <- subset(bcicensusdat, !is.na(dbh) & production67 > 0 & comp_idx > 0)
twoslopeplot(dat = alltreedat, 
             plottitle = 'All species', 
             xl = xl1, 
             yl =  yl1,
             binvar = 'mass')

shadedat <- subset(bcicensusdat, !is.na(dbh) & production67 > 0 & comp_idx > 0 & tol_wright == 'S')
twoslopeplot(dat = shadedat, 
             plottitle = 'Shade-tolerant species', 
             xl = xl1, 
             yl =  yl1,
             binvar = 'mass')

intdat <- subset(bcicensusdat, !is.na(dbh) & production67 > 0 & comp_idx > 0 & tol_wright == 'I')
twoslopeplot(dat = intdat, 
             plottitle = 'Intermediate species', 
             xl = xl1, 
             yl =  yl1,
             binvar = 'mass')

gapdat <- subset(bcicensusdat, !is.na(dbh) & production67 > 0 & comp_idx > 0 & tol_wright == 'G')
twoslopeplot(dat = gapdat, 
             plottitle = 'Gap species', 
             xl = xl1, 
             yl =  yl1,
             binvar = 'mass')
```

\newpage
## Binned density

```{r, results='hide', fig.keep = 'none', warning = FALSE, message = FALSE}
numbins <- 20 # Can be edited if desired. Only for visualization purposes.

# Pareto fits
bci_dens_fit_all <- powerlawfit(alltreedat$agb)
bci_dens_fit_shade <- powerlawfit(shadedat$agb)
bci_dens_fit_inter <- powerlawfit(intdat$agb)
bci_dens_fit_gap <- powerlawfit(gapdat$agb)

# Log bins
bci_dens_logbin_all <- with(alltreedat, logbin(x=agb, y=NULL, n = numbins))
bci_dens_logbin_shade <- with(shadedat, logbin(x=agb, y=NULL, n = numbins))
bci_dens_logbin_inter <- with(intdat, logbin(x=agb, y=NULL, n = numbins))
bci_dens_logbin_gap <- with(gapdat, logbin(x=agb, y=NULL, n = numbins))
```

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.width = 3.25, fig.height = 3.25}
xl2 <- 'Mass (kg)'
yl2 <- 'Density'

plotbinsandfits(bci_dens_fit_all, bci_dens_logbin_all,
                plottitle = 'All species', xl = xl2, yl = yl2)

plotbinsandfits(bci_dens_fit_shade, bci_dens_logbin_shade,
                plottitle = 'Shade-tolerant species', xl = xl2, yl = yl2)

plotbinsandfits(bci_dens_fit_inter, bci_dens_logbin_inter,
                plottitle = 'Intermediate species', xl = xl2, yl = yl2)

plotbinsandfits(bci_dens_fit_gap, bci_dens_logbin_gap,
                plottitle = 'Gap species', xl = xl2, yl = yl2)
```

\newpage
## Binned production

```{r, results = 'hide', fig.keep = 'none'}
numbins <- 20 # Can be edited if desired. ***NOT JUST FOR LOOKS***

# Log bins
bci_prod_logbin_all <- with(alltreedat, 
                   logbin(x=agb, y=production67, n=numbins))
bci_prod_logbin_shade <- with(shadedat, 
                         logbin(x=agb, y=production67, n=numbins))
bci_prod_logbin_inter <- with(intdat, 
                         logbin(x=agb, y=production67, n=numbins))
bci_prod_logbin_gap <- with(gapdat, 
                       logbin(x=agb, y=production67, n=numbins))

# Energy-equivalence slopes
bci_prod_lm_all <- lm(log10(bin_value) ~ log10(bin_midpoint), data = bci_prod_logbin_all)
bci_prod_lm_shade <- lm(log10(bin_value) ~ log10(bin_midpoint), data = bci_prod_logbin_shade)
bci_prod_lm_inter <- lm(log10(bin_value) ~ log10(bin_midpoint), data = bci_prod_logbin_inter)
bci_prod_lm_gap <- lm(log10(bin_value) ~ log10(bin_midpoint), data = bci_prod_logbin_gap)

```

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.width = 3.25, fig.height = 3.25}
xl3 <- 'Mass (kg)'
yl3 <- expression(paste('Total production (kg y'^-1,')', sep=''))

plotlogbin(bci_prod_logbin_all, xl3, yl3, 
           'All species', reg = TRUE)
plotlogbin(bci_prod_logbin_shade, xl3, yl3,  
           'Shade-tolerant species', reg = TRUE)
plotlogbin(bci_prod_logbin_inter, xl3, yl3, 
           'Intermediate species', reg = TRUE)
plotlogbin(bci_prod_logbin_gap, xl3, yl3, 
           'Gap species', reg = TRUE)
```

\newpage
## Slope estimates, with confidence intervals

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.width = 5, fig.height = 3.25}

lmsize_all <- lm(log10(production67) ~ log10(agb), data=alltreedat)
lmsizecomp_all <- lm(log10(production67) ~ log10(agb) + log10(comp_idx), data=alltreedat)
lmsize_shade <- lm(log10(production67) ~ log10(agb), data=shadedat)
lmsizecomp_shade <- lm(log10(production67) ~ log10(agb) + log10(comp_idx), data=shadedat)
lmsize_int <- lm(log10(production67) ~ log10(agb), data=intdat)
lmsizecomp_int <- lm(log10(production67) ~ log10(agb) + log10(comp_idx), data=intdat)
lmsize_gap <- lm(log10(production67) ~ log10(agb), data=gapdat)
lmsizecomp_gap <- lm(log10(production67) ~ log10(agb) + log10(comp_idx), data=gapdat)

slopedat <- data.frame(withcompidx = rep(c('comp idx','no comp idx'), each=4),
                       guild = c('all','shade','intermediate','gap'),
                       slope = c(lmsizecomp_all$coef[2], lmsizecomp_shade$coef[2], lmsizecomp_int$coef[2], lmsizecomp_gap$coef[2],
                                 lmsize_all$coef[2], lmsize_shade$coef[2], lmsize_int$coef[2], lmsize_gap$coef[2]),
                       cimin = c(confint(lmsizecomp_all)[2,1], confint(lmsizecomp_shade)[2,1], confint(lmsizecomp_int)[2,1], confint(lmsizecomp_gap)[2,1],
                                 confint(lmsize_all)[2,1], confint(lmsize_shade)[2,1], confint(lmsize_int)[2,1], confint(lmsize_gap)[2,1]),
                       cimax = c(confint(lmsizecomp_all)[2,2], confint(lmsizecomp_shade)[2,2], confint(lmsizecomp_int)[2,2], confint(lmsizecomp_gap)[2,2],
                                 confint(lmsize_all)[2,2], confint(lmsize_shade)[2,2], confint(lmsize_int)[2,2], confint(lmsize_gap)[2,2]))

ggplot(slopedat, aes(x = guild, y = slope, ymin = cimin, ymax = cimax)) +
  facet_grid(. ~ withcompidx) +
  geom_pointrange() +
  panel_border(colour = 'black') +
  ggtitle('Individual production')
```

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.width = 3.25, fig.height = 3.25}
# Density
slopedat <- data.frame(guild = c('all','shade','intermediate','gap'),
                       slope = c(bci_dens_fit_all$alpha, bci_dens_fit_shade$alpha, bci_dens_fit_inter$alpha, bci_dens_fit_gap$alpha),
                       cimin = c(bci_dens_fit_all$boot_ci[1], bci_dens_fit_shade$boot_ci[1], bci_dens_fit_inter$boot_ci[1], bci_dens_fit_gap$boot_ci[1]),
                       cimax = c(bci_dens_fit_all$boot_ci[2], bci_dens_fit_shade$boot_ci[2], bci_dens_fit_inter$boot_ci[2], bci_dens_fit_gap$boot_ci[2]))

ggplot(slopedat, aes(x = guild, y = slope, ymin = cimin, ymax = cimax)) +
  geom_pointrange() +
  panel_border(colour = 'black') +
  ggtitle('Density by mass')

# Binned production
slopedat <- data.frame(guild = c('all','shade','intermediate','gap'),
                       slope = c(bci_prod_lm_all$coef[2], bci_prod_lm_shade$coef[2], bci_prod_lm_inter$coef[2], bci_prod_lm_gap$coef[2]),
                       cimin = c(confint(bci_prod_lm_all)[2,1], confint(bci_prod_lm_shade)[2,1], confint(bci_prod_lm_inter)[2,1], confint(bci_prod_lm_gap)[2,1]),
                       cimax = c(confint(bci_prod_lm_all)[2,2], confint(bci_prod_lm_shade)[2,2], confint(bci_prod_lm_inter)[2,2], confint(bci_prod_lm_gap)[2,2]))

ggplot(slopedat, aes(x = guild, y = slope, ymin = cimin, ymax = cimax)) +
  geom_pointrange() +
  panel_border(colour = 'black') +
  ggtitle('Production binned by mass')

```

Guild            | Density | Indiv. production | Density + indiv. production | Binned production | Indiv. production with comp. index
---------------- | ------- | ----------------- | --------------------------- | ----------------- | ----------------------------------
All trees        | -1.26   | 0.78              | -0.48                       | -0.59             | 0.59
Shade-tolerant   | -1.33   | 0.76              | -0.57                       | -0.70             | 0.73
Intermediate     | -1.28   | 0.78              | -0.50                       | -0.51             | 0.72
Gap/pioneer      | -1.34   | 0.81              | -0.53                       | -0.52             | 0.78

