---
title: "Density and production scaling relationships"
author: "Quentin D. Read"
date: "March 21, 2017"
output: 
  pdf_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

## What I did

Below, I used the *poweRlaw* package to fit continuous Pareto (exponential) and lognormal distributions to density scaling relationships for BCI and Harvard Forest. I split each site into shade guilds and fit the same distributions. I plotted the two fits. 

In addition, I coded Ethan White's log-binning algorithm and binned production by biomass for BCI and Harvard. As before, I did this for all species, and for the three shade guilds at each site. I fit a log-linear regression to these bins. Finally, I did the individual production by biomass scaling relationships as a log-linear fit.

Overall, it looks like the lognormal is a better fit in almost every case than the Pareto. It looks like the deviations from the lognormal at high biomasses match the results that Caroline Farrior got for the density scaling relationships. There might be some interesting differences in the fits between shade guilds for both the density scaling and the production scaling relationships. Also, the binned production plots do not seem to support the predictions of energy equivalence.

## What I will do

For now, I only estimated the $\alpha$ parameter for the Pareto distributions, and the $\mu$ and $\sigma$ parameters for the lognormal distributions. Each of them also has an $x_{min}$ parameter that can be estimated from the data, but for now, I just used the minimum value in the data as the $x_{min}$ parameter. I think that is appropriate but I can also try to estimate it from the data.

Also, there are no confidence intervals on the parameter estimates because I have not yet run the bootstraps. They will be added soon.

## Version history

* 21 March: all relevant plots are changed to histograms, and the individual production scaling is added. There is also a table at the bottom of the document.
* 20 March: changed plots to PDF instead of CDF, edited axis ticks.
* 17 March: changed axis labels and made y-axes logarithmic
* 16 March: created document

\newpage

# Methods

## Preparing data

The biomass data are year 3. The production data are the increment of biomass between years 1 and 3, halved.

```{r, message=FALSE}
source('~/GitHub/forestlight/code/mergeshade.r')

pdat <- dat %>% 
  mutate(biomass3 = pmax(BiomassDGH_year3_kg, BiomassDBH_year3_kg, na.rm=TRUE),
         biomass2 = pmax(BiomassDGH_year2_kg, BiomassDBH_year2_kg, 0, na.rm=TRUE),
         biomass1 = pmax(BiomassDGH_year1_kg, BiomassDBH_year1_kg, 0, na.rm=TRUE),
         massprod23 = pmax(biomass3 - biomass2, 0, na.rm=TRUE),
         massprod12 = pmax(biomass2 - biomass1, 0, na.rm=TRUE),
         massprod13 = pmax((biomass3 - biomass1)/2, 0, na.rm=TRUE)) %>% 
  select(Site, biomass3, biomass2, massprod23, massprod12, massprod13, CII, tolerance)
bcidat <- pdat %>% filter(Site == 'Barro_Colorado_Island', !is.na(CII), !is.na(biomass3))
harvdat <- pdat %>% filter(Site == 'Harvard_Forest_LTER', !is.na(CII), !is.na(biomass3))
```

[Click to return to table of contents](#top)

\newpage

## Functions needed to run analysis

This function fits both continuous Pareto and continuous lognormal to the data (input as a single vector of values), and returns their parameters plus all data needed to plot them. This returns the pdf as well.

```{r}
powerlawfit <- function(dat) {
  library(poweRlaw)
  pl_dat <- conpl$new(dat)
  lognorm_dat <- conlnorm$new(dat)
  xmin_pl <- pl_dat$getXmin()
  xmin_lognorm <- lognorm_dat$getXmin()
  pars_pl <- estimate_pars(pl_dat)
  pars_lognorm <- estimate_pars(lognorm_dat)
  pl_dat$setPars(pars_pl)
  lognorm_dat$setPars(pars_lognorm)
  plotdat <- plot(pl_dat)
  plfit_dat <- lines(pl_dat)
  lognormfit_dat <- lines(lognorm_dat)
  pl_pdf <- dist_pdf(m = pl_dat, q = plfit_dat$x, log = FALSE)
  lognorm_pdf <- dist_pdf(m = lognorm_dat, q = lognormfit_dat$x, log = FALSE)
  
  return(list(plotdat = plotdat, 
              plfit = plfit_dat, 
              lognormfit = lognormfit_dat, 
              plpdf = data.frame(x = plfit_dat$x, y = pl_pdf),
              lognormpdf = data.frame(x = lognormfit_dat$x, y = lognorm_pdf),
              xmin = xmin_pl, 
              alpha = pars_pl$pars,
              xmin_lognorm = xmin_lognorm,
              pars_lognorm = pars_lognorm$pars))
}
```

\newpage
This function plots the data in a way that I think is nice.

```{r}
plotpowerlawfit <- function(pl, xl, title, subtitle) {
  library(ggplot2)
  th <- theme_bw() + theme(panel.grid = element_blank())
  
  expr1 <- as.character(as.expression(substitute(
    "Pareto:"~~alpha == a, list(a = round(pl$alpha, 2)))))
  expr2 <- as.character(as.expression(substitute(
    "Lognormal:"~~mu == m*","~~sigma == s, list(m = round(pl$pars_lognorm[1], 2), 
                                                s = round(pl$pars_lognorm[2], 2)))))

  p <- ggplot(pl$plotdat, aes(x = x, y = y)) +
    scale_x_log10(breaks = scales::trans_breaks("log10", function(x) 10^x),
                  labels = scales::trans_format("log10", scales::math_format(10^.x))) +
    scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x),
                  labels = scales::trans_format("log10", scales::math_format(10^.x))) +
    geom_point() +
    geom_line(data = subset(pl$plfit, y>0), color = 'forestgreen') +
    geom_line(data = subset(pl$lognormfit, y>0), color = 'indianred') +
    geom_text(x = -Inf, y = -Inf, label = expr1, parse = TRUE, hjust = 0, vjust = -1.5) +
    geom_text(x = -Inf, y = -Inf, label = expr2, parse = TRUE, hjust = 0, vjust = -0.25) +
    th +
    labs(x = xl, y = 'log CDF') +
    ggtitle(title, subtitle)
  return(p)
}
```

\newpage
This function implements Ethan White's log-binning algorithm. For our purposes, $x$ is biomass, $y$ is production, and $n$ is the desired number of bins. Modified on 20 March to also include 1-dimensional log-binning algorithm (for the density scaling bins).

```{r}
logbin <- function(x, y = NULL, n) {
  logx <- log10(x)                                           # log transform x value (biomass)
  bin_edges <- seq(min(logx), max(logx), length.out = n + 1) # get edges of bins
  logxbin <- rep(NA, length(logx))                           # create data structure to assign trees to bins
  b <- bin_edges                                             # add a little to the biggest bin temporarily
  b[length(b)] <- b[length(b)] + 1                           # (so that the biggest single tree is put in a bin)
  for (i in 1:length(logx)) {
    logxbin[i] <- sum(logx[i] >= b)                          # assign each tree to a bin
  }
  bin_midpoints <- numeric(n)
  for (i in 1:n) {
    bin_midpoints[i] <- mean(10^(bin_edges[i:(i+1)]))        # backtransform bin edges to linear, and get midpoints
  }
  bin_widths <- diff(10^bin_edges)                           # get linear width of each bin
  bin_factor <- factor(logxbin, levels=1:n)                  # convert bin to factor (required to deal with zeroes if present)
  bin_counts <- table(bin_factor)                            # find number of trees in each bin
  if (!is.null(y)) {
    rawy <- tapply(y, bin_factor, sum)                       # sum y value (production) in each bin
    rawy[is.na(rawy)] <- 0                                   # add zeroes back in if present
    bin_values <- as.numeric(rawy/bin_widths)                # divide production by width for each bin 
  }
  else {
    bin_values <- as.numeric(bin_counts/bin_widths)          # 1-dimensional case.
  }
  
  return(data.frame(bin_midpoint = bin_midpoints,            # return result!
                    bin_value = bin_values,                  # also add bin min and max for bar plot purposes
                    bin_count = as.numeric(bin_counts),
                    bin_min = 10^bin_edges[1:n],
                    bin_max = 10^bin_edges[2:(n+1)]))
  
}
```

This is a verbal description of the binning algorithm given in R code above.

1. Log transform the year 3 biomass data
2. Bin into equal width bins (on the logarithmic axis)
3. Back transform the bin edges to linear axis.
4. Get the midpoint of each bin. This is the X-value on the plot.
5. Sum the biomass production of each bin, expressed as (year 3 biomass - year 1 biomass)/2
6. Divide each bin's summed production value by the linear width of each bin (this will be the Y-value on the plot).

\newpage
This function plots the log bins as a histogram. Both x and y-axes are logarithmic. It can also plot a log-linear regression fit.

```{r}
plotlogbin <- function(dat, xl, yl, title, subtitle, reg = FALSE) {
  library(ggplot2)
  th <- theme_bw() + theme(panel.grid = element_blank())
  
  y_min <- 10^floor(log10(min(dat$bin_value, na.rm = TRUE)))
  y_max <- max(dat$bin_value, na.rm = TRUE) * 1.1
  
  p <- ggplot(dat, aes(xmin=bin_min, xmax=bin_max, ymin=0, ymax=bin_value)) + 
    geom_rect(alpha = 0.5) +
    scale_x_log10(name = xl, expand = c(0,0),
                  breaks = scales::trans_breaks("log10", function(x) 10^x),
                  labels = scales::trans_format("log10", scales::math_format(10^.x))) +
    scale_y_log10(name = yl, expand = c(0,0), limits = c(y_min, y_max),
                  breaks = scales::trans_breaks("log10", function(x) 10^x),
                  labels = scales::trans_format("log10", scales::math_format(10^.x))) +
    th + ggtitle(title, subtitle)
  if (reg) {
    p <- p +
      stat_smooth(method = 'lm', se = FALSE, color = 'black', 
                  aes(x = bin_midpoint, y = bin_value)) +
      geom_text(x = -Inf, y = -Inf, 
                label = paste('Slope:', round(lm(I(log10(bin_value)) ~ I(log10(bin_midpoint)), data=dat)$coef[2], 2)),
                hjust = 0, vjust = -0.25)
  }
  return(p)
}
```

\newpage
This function, added on 20 March, plots the PDFs (not CDFs) of the Pareto and lognormal fits on a histogram of the data. It is necessary to make bins because that is the only way to plot the raw data behind the PDF.

```{r}
plotbinsandfits <- function(pl, bindat, xl, yl = 'log PDF', title, subtitle) {
  library(ggplot2)
  th <- theme_bw() + theme(panel.grid = element_blank())
  
  expr1 <- as.character(as.expression(substitute(
    "Pareto:"~~alpha == a, list(a = round(pl$alpha, 2)))))
  expr2 <- as.character(as.expression(substitute(
    "Lognormal:"~~mu == m*","~~sigma == s, list(m = round(pl$pars_lognorm[1], 2), 
                                                s = round(pl$pars_lognorm[2], 2)))))
  
  bindat <- transform(bindat, bin_value = bin_value/sum(bin_count))
  bindat <- subset(bindat, bin_value > 0)
  y_min <- 10^floor(log10(min(bindat$bin_value, na.rm = TRUE)))
  y_max <- max(bindat$bin_value, na.rm = TRUE) * 1.1
  
  p <- ggplot(bindat) + 
    scale_x_log10(name = xl,
                  breaks = scales::trans_breaks("log10", function(x) 10^x),
                  labels = scales::trans_format("log10", scales::math_format(10^.x)),
                  expand = c(0,0)) +
    scale_y_log10(name = yl,
                  limits = c(y_min, y_max),
                  breaks = scales::trans_breaks("log10", function(x) 10^x),
                  labels = scales::trans_format("log10", scales::math_format(10^.x)),
                  expand=c(0,0)) +
    geom_rect(aes(xmin = bin_min, xmax = bin_max, ymin = 0, ymax = bin_value), alpha = 0.5) +
    geom_line(data = subset(pl$plpdf, y>0), aes(x,y), color = 'forestgreen') +
    geom_line(data = subset(pl$lognormpdf, y>0), aes(x,y), color = 'indianred') +
    geom_text(x = -Inf, y = -Inf, label = expr1, parse = TRUE, hjust = 0, vjust = -1.5) +
    geom_text(x = -Inf, y = -Inf, label = expr2, parse = TRUE, hjust = 0, vjust = -0.25) +
    th +
    labs(x = xl, y = 'log PDF') +
    ggtitle(title, subtitle)
  return(p)
}
```


[Click to return to table of contents](#top)

\newpage

# Results

## Density scaling relationships

Below, I fit all the power laws at once, then plot the results. **Note: For all plots below, the green line is the Pareto fit and the red line is the lognormal fit.** Note here that the Pareto $\alpha$ parameter is equivalent to -slope, so, for example, $\alpha = 1.35$ corresponds to a slope of -1.35.

```{r, results='hide', fig.keep = 'none', warning = FALSE}
numbins <- 20

bci_abund_all <- powerlawfit(bcidat$biomass3)
bci_abund_shade <- powerlawfit(subset(bcidat, tolerance == 'S')$biomass3)
bci_abund_inter <- powerlawfit(subset(bcidat, tolerance == 'I')$biomass3)
bci_abund_gap <- powerlawfit(subset(bcidat, tolerance == 'G')$biomass3)
harv_abund_all <- powerlawfit(harvdat$biomass3)
harv_abund_shade <- powerlawfit(subset(harvdat, tolerance == 'S')$biomass3)
harv_abund_inter <- powerlawfit(subset(harvdat, tolerance == 'I')$biomass3)
harv_abund_gap <- powerlawfit(subset(harvdat, tolerance == 'G')$biomass3)
bci_logbin_all <- with(bcidat, logbin(x=biomass3, y=NULL, n = numbins))
bci_logbin_shade <- with(subset(bcidat, tolerance == 'S'), logbin(x=biomass3, y=NULL, n = numbins))
bci_logbin_inter <- with(subset(bcidat, tolerance == 'I'), logbin(x=biomass3, y=NULL, n = 15))
bci_logbin_gap <- with(subset(bcidat, tolerance == 'G'), logbin(x=biomass3, y=NULL, n = 15))
harv_logbin_all <- with(harvdat, logbin(x=biomass3, y=NULL, n = numbins))
harv_logbin_shade <- with(subset(harvdat, tolerance == 'S'), logbin(x=biomass3, y=NULL, n = numbins))
harv_logbin_inter <- with(subset(harvdat, tolerance == 'I'), logbin(x=biomass3, y=NULL, n = 10))
harv_logbin_gap <- with(subset(harvdat, tolerance == 'G'), logbin(x=biomass3, y=NULL, n = 8))

```

\newpage

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.width = 3.25, fig.height = 3.25}
label1 <- 'Individual mass (kg)'

plotbinsandfits(bci_abund_all, bci_logbin_all, label1, 'log PDF', 'Density scaling, BCI', 'all species')
plotbinsandfits(bci_abund_shade, bci_logbin_shade, label1, 'log PDF', 'Density scaling, BCI', 'shade-tolerant species')
plotbinsandfits(bci_abund_inter, bci_logbin_inter, label1,'log PDF', 'Density scaling, BCI', 'intermediate species')
plotbinsandfits(bci_abund_gap, bci_logbin_gap, label1,'log PDF', 'Density scaling, BCI', 'gap species')
plotbinsandfits(harv_abund_all, harv_logbin_all, label1, 'log PDF', 'Density scaling, Harvard', 'all species')
plotbinsandfits(harv_abund_shade, harv_logbin_shade, label1, 'log PDF', 'Density scaling, Harvard', 'shade-tolerant species')
plotbinsandfits(harv_abund_inter, harv_logbin_inter, label1, 'log PDF', 'Density scaling, Harvard', 'intermediate species')
plotbinsandfits(harv_abund_gap, harv_logbin_gap, label1, 'log PDF', 'Density scaling, Harvard', 'gap species')
```

[Click to return to table of contents](#top)

\newpage

## Individual production by individual mass scaling relationships

This has been modified on 21 March to use the simpler method of fitting a regression to the log-transformed production and biomass values for each tree. However, the major caveat to this set of relationships is that the power law cannot be fit to 0 values. The 0 values probably represent trees with very low growth, but it is impossible to say. Therefore, I had to throw them out, but they comprise a pretty big portion of the trees (23% of trees in BCI and 20% of trees in HF).

```{r}
bci_prod_all <- lm(log10(massprod13) ~ log10(biomass3), data = bcidat, subset = massprod13 > 0)
bci_prod_shade <- lm(log10(massprod13) ~ log10(biomass3), data = bcidat, subset = massprod13 > 0 & tolerance == 'S')
bci_prod_inter <- lm(log10(massprod13) ~ log10(biomass3), data = bcidat, subset = massprod13 > 0 & tolerance == 'I')
bci_prod_gap <- lm(log10(massprod13) ~ log10(biomass3), data = bcidat, subset = massprod13 > 0 & tolerance == 'G')
harv_prod_all <- lm(log10(massprod13) ~ log10(biomass3), data = harvdat, subset = massprod13 > 0)
harv_prod_shade <- lm(log10(massprod13) ~ log10(biomass3), data = harvdat, subset = massprod13 > 0 & tolerance == 'S')
harv_prod_inter <- lm(log10(massprod13) ~ log10(biomass3), data = harvdat, subset = massprod13 > 0 & tolerance == 'I')
harv_prod_gap <- lm(log10(massprod13) ~ log10(biomass3), data = harvdat, subset = massprod13 > 0 & tolerance == 'G')
```

\newpage

For each site, I plotted each of the regression relationships by guild. Here, I have excluded the points with no tolerance score from displaying on the plot. However, the overall regression line does include those points in its calculation.

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.width = 3.25, fig.height = 3.25}

bcislopes <- data.frame(biomass3 = -Inf, 
                        massprod13 = Inf, 
                        txt = c(paste('All trees:',round(bci_prod_all$coef[2], 2)),
                                paste('Shade-tolerant:',round(bci_prod_shade$coef[2], 2)),
                                paste('Intermediate:',round(bci_prod_inter$coef[2], 2)),
                                paste('Gap:',round(bci_prod_gap$coef[2], 2))
                                ))
harslopes <- data.frame(biomass3 = -Inf, 
                        massprod13 = Inf, 
                        txt = c(paste('All trees:',round(harv_prod_all$coef[2], 2)),
                                paste('Shade-tolerant:',round(harv_prod_shade$coef[2], 2)),
                                paste('Intermediate:',round(harv_prod_inter$coef[2], 2)),
                                paste('Gap:',round(harv_prod_gap$coef[2], 2))
                                ))

ggplot(subset(bcidat, massprod13 > 0), aes(x = biomass3, y = massprod13)) +
  geom_point(aes(color = tolerance), alpha = 0.5, data = subset(bcidat, massprod13 > 0 & !is.na(tolerance))) +
  stat_smooth(aes(group = tolerance, color = tolerance), method = 'lm', se = FALSE, data = subset(bcidat, massprod13 > 0 & !is.na(tolerance))) +
  stat_smooth(method = 'lm', se = FALSE, color = 'black') +
  scale_x_log10(name = 'Individual mass (kg)',
                breaks = scales::trans_breaks("log10", function(x) 10^x),
                labels = scales::trans_format("log10", scales::math_format(10^.x)),
                expand = c(0,0)) +
  scale_y_log10(name = expression(paste('Individual production (kg y'^-1,')', sep = '')),
                breaks = scales::trans_breaks("log10", function(x) 10^x),
                labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  geom_text(x = -Inf, y = Inf, label = bcislopes$txt[1], hjust = 0, vjust = 1) +
  geom_text(x = -Inf, y = Inf, label = bcislopes$txt[2], hjust = 0, vjust = 2) +
  geom_text(x = -Inf, y = Inf, label = bcislopes$txt[3], hjust = 0, vjust = 3) +
  geom_text(x = -Inf, y = Inf, label = bcislopes$txt[4], hjust = 0, vjust = 4) +
  theme_bw() + theme(panel.grid = element_blank()) +
  ggtitle('Individual production scaling', 'BCI')

ggplot(subset(harvdat, massprod13 > 0), aes(x = biomass3, y = massprod13)) +
  geom_point(aes(color = tolerance), alpha = 0.5, data = subset(harvdat, massprod13 > 0 & !is.na(tolerance))) +
  stat_smooth(aes(group = tolerance, color = tolerance), method = 'lm', se = FALSE, data = subset(harvdat, massprod13 > 0 & !is.na(tolerance))) +
  stat_smooth(method = 'lm', se = FALSE, color = 'black') +
  scale_x_log10(name = 'Individual mass (kg)',
                breaks = scales::trans_breaks("log10", function(x) 10^x),
                labels = scales::trans_format("log10", scales::math_format(10^.x)),
                expand = c(0,0)) +
  scale_y_log10(name = expression(paste('Individual production (kg y'^-1,')', sep = '')),
                breaks = scales::trans_breaks("log10", function(x) 10^x),
                labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  geom_text(x = -Inf, y = Inf, label = harslopes$txt[1], hjust = 0, vjust = 1) +
  geom_text(x = -Inf, y = Inf, label = harslopes$txt[2], hjust = 0, vjust = 2) +
  geom_text(x = -Inf, y = Inf, label = harslopes$txt[3], hjust = 0, vjust = 3) +
  geom_text(x = -Inf, y = Inf, label = harslopes$txt[4], hjust = 0, vjust = 4) +
  theme_bw() + theme(panel.grid = element_blank()) +
  ggtitle('Individual production scaling', 'Harvard Forest')
```

[Click to return to table of contents](#top)

\newpage

## Total binned production scaling relationships

For binning, I tried to create 20 bins if possible. Then, I simply fit a regression line to the relationship between $\log_{10} production$ and $\log_{10} mass$ which is the slope that we should use to evaluate the predictions of EE. Previously I tried to do a power law fit after binning, but that was dumb because power law fit is only for unbinned data!

```{r, results = 'hide', fig.keep = 'none'}
numbins <- 20

bci_logbin <- with(bcidat, logbin(x=biomass3, y=massprod13, n=numbins))
bci_logbin_shade <- with(subset(bcidat, tolerance == 'S'), logbin(x=biomass3, y=massprod13, n=numbins))
bci_logbin_inter <- with(subset(bcidat, tolerance == 'I'), logbin(x=biomass3, y=massprod13, n=15))
bci_logbin_gap <- with(subset(bcidat, tolerance == 'G'), logbin(x=biomass3, y=massprod13, n=15))
harv_logbin <- with(harvdat, logbin(x=biomass3, y=massprod13, n=numbins))
harv_logbin_shade <- with(subset(harvdat, tolerance == 'S'), logbin(x=biomass3, y=massprod13, n=numbins))
harv_logbin_inter <- with(subset(harvdat, tolerance == 'I'), logbin(x=biomass3, y=massprod13, n=10))
harv_logbin_gap <- with(subset(harvdat, tolerance == 'G'), logbin(x=biomass3, y=massprod13, n=8))

bci_lm_all <- lm(log10(bin_value) ~ log10(bin_midpoint), data = bci_logbin)
bci_lm_shade <- lm(log10(bin_value) ~ log10(bin_midpoint), data = bci_logbin_shade)
bci_lm_inter <- lm(log10(bin_value) ~ log10(bin_midpoint), data = bci_logbin_inter)
bci_lm_gap <- lm(log10(bin_value) ~ log10(bin_midpoint), data = bci_logbin_gap)
harv_lm_all <- lm(log10(bin_value) ~ log10(bin_midpoint), data = harv_logbin)
harv_lm_shade <- lm(log10(bin_value) ~ log10(bin_midpoint), data = harv_logbin_shade)
harv_lm_inter <- lm(log10(bin_value) ~ log10(bin_midpoint), data = harv_logbin_inter)
harv_lm_gap <- lm(log10(bin_value) ~ log10(bin_midpoint), data = harv_logbin_gap)

```

\newpage

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.width = 3.25, fig.height = 3.25}
label2 <- expression(paste('Total production (kg y'^-1,')', sep = ''))

plotlogbin(bci_logbin, label1, label2, 'Production bins, BCI', 'all species', reg = TRUE)
plotlogbin(bci_logbin_shade, label1, label2, 'Production bins, BCI', 'shade-tolerant species', reg = TRUE)
plotlogbin(bci_logbin_inter, label1, label2, 'Production bins, BCI', 'intermediate species', reg = TRUE)
plotlogbin(bci_logbin_gap, label1, label2, 'Production bins, BCI', 'gap species', reg = TRUE)
plotlogbin(harv_logbin, label1, label2, 'Production bins, Harvard', 'all species', reg = TRUE)
plotlogbin(harv_logbin_shade, label1, label2, 'Production bins, Harvard', 'shade-tolerant species', reg = TRUE)
plotlogbin(harv_logbin_inter, label1, label2, 'Production bins, Harvard', 'intermediate species', reg = TRUE)
plotlogbin(harv_logbin_gap, label1, label2, 'Production bins, Harvard', 'gap species', reg = TRUE)
```

[Click to return to table of contents](#top)

\newpage

## Table of scaling coefficients

The third coefficient column is the sum of the first two. The density column is taken from the Pareto regression, while individual and binned production are from log-linear regressions.

Site               | Guild            | Density | Indiv. production | Density + indiv. production | Binned production
------------------ | ---------------- | ------- | ----------------- | --------------------------- | -----------------
Barro Colorado     | All trees        | -1.35   | 0.79              | -0.56                       | -0.66
Barro Colorado     | Shade-tolerant   | -1.33   | 0.86              | -0.47                       | -0.60
Barro Colorado     | Intermediate     | -1.35   | 0.92              | -0.43                       | -0.37
Barro Colorado     | Gap/pioneer      | -1.26   | 0.74              | -0.52                       | -0.44
Harvard Forest     | All trees        | -1.22   | 0.70              | -0.52                       | -0.33
Harvard Forest     | Shade-tolerant   | -1.25   | 0.70              | -0.55                       | -0.37
Harvard Forest     | Intermediate     | -1.14   | 0.82              | -0.32                       | 0.20
Harvard Forest     | Gap/pioneer      | -1.19   | 0.70              | -0.49                       | -0.02

[Click to return to table of contents](#top)