---
title: "Ratio of two power laws"
author: "Quentin D. Read"
date: "February 7, 2019"
output: pdf_document
---

# Derivation

The Pareto probability density function of abundance in functional group 1 as a function of diameter is:

$$ y_{1} = \frac { \alpha_{1} x_{min1}^{\alpha_{1}} } { x^{\alpha_{1} + 1} } $$

Take the logarithm of the ratio of the density functions of two functional groups:

$$ \log \frac {y_{1}}{y_{2}} = \log \frac {\frac { \alpha_{1} x_{min1}^{\alpha_{1}} } { x^{\alpha_{1} + 1} }} {\frac { \alpha_{2} x_{min2}^{\alpha_{2}} } { x^{\alpha_{2} + 1} }} $$
Simplify:

$$ \log \frac {y_{1}}{y_{2}} = \log \frac {\alpha_{1} x_{min1}^{\alpha_{1}}} {\alpha_{2} x_{min2}^{\alpha_{2}}} + \log {x ^ {\alpha_{2} - \alpha_{1}} } $$
$$ \log \frac {y_{1}}{y_{2}} = \log \frac {\alpha_{1} x_{min1}^{\alpha_{1}}} {\alpha_{2} x_{min2}^{\alpha_{2}}} + ({\alpha_{2} - \alpha_{1}}) \log x  $$
The first term is a constant so the slope of the ratio of the two functional groups in log space is ${\alpha_{2} - \alpha_{1}}$. Obviously this slope could also be calculated piecewise if the two functional groups have piecewise powerlaw fits, or the scaling variable $x$ could represent light per area instead of diameter.

# Example

Generate 100000 random draws from two Pareto distributions one with parameter $\alpha_{1}=1$ (slope of -2) and the other with $\alpha_{2}=2$ (slope of -3), and both with $x_{min}=1$. The ratio of the bins should have slope 1.

```{r, message=FALSE, warning=FALSE}
library(actuar)
n <- 100000
alpha <- c(1, 2)
xmin <- 1

set.seed(111)
y1 <- rpareto1(n = n, shape = alpha[1], min = xmin)
y2 <- rpareto1(n = n, shape = alpha[2], min = xmin)
```

This is the log bin function, described in great detail elsewhere.

```{r}
logbin_setedges <- function(x, y = NULL, edges) {
  logx <- log10(x)                                        
  bin_edges <- log10(c(edges$bin_min, edges$bin_max[length(edges$bin_max)]))
  n <- length(edges$bin_min)
  logxbin <- rep(NA, length(logx))                          
  b <- bin_edges                                            
  b[length(b)] <- b[length(b)] + 1                          
  for (i in 1:length(logx)) {
    logxbin[i] <- sum(logx[i] >= b)                          
  }
  bin_midpoints <- edges$bin_midpoint
  bin_widths <- diff(10^bin_edges)                          
  bin_factor <- factor(logxbin, levels=1:n)                  
  bin_counts <- table(bin_factor)                           
  if (!is.null(y)) {
    rawy <- tapply(y, bin_factor, sum)                      
    rawy[is.na(rawy)] <- 0                                   
    bin_values <- as.numeric(rawy/bin_widths)                
  }
  else {
    bin_values <- as.numeric(bin_counts/bin_widths)         
  }
  
  return(data.frame(bin_midpoint = bin_midpoints,            
                    bin_value = bin_values,                  
                    bin_count = as.numeric(bin_counts),
                    bin_min = 10^bin_edges[1:n],
                    bin_max = 10^bin_edges[2:(n+1)]))
  
}
```

Calculate the log bins.

```{r}
n_bins <- 10
bin_bounds <- exp(seq(log(1), log(min(max(y1), max(y2))), length.out = n_bins + 1))
bin_data <- data.frame(bin_min = bin_bounds[1:n_bins], 
                       bin_max = bin_bounds[2:length(bin_bounds)],
                       bin_midpoint = exp(diff(log(bin_bounds))/2 + log(bin_bounds[1:n_bins])))

y1bins <- logbin_setedges(y1, edges = bin_data)
y2bins <- logbin_setedges(y2, edges = bin_data)
```

\newpage

Plot the log bins with a reference line of slope 1. Note that the intercept of the reference line is equal to $\log \frac {\alpha_{1} x_{min1}^{\alpha_{1}}} {\alpha_{2} x_{min2}^{\alpha_{2}}}$ as in the equation above. Aspect ratio is set to 1:1.

```{r, message = FALSE}
library(ggplot2)

plotdat <- data.frame(x = y1bins$bin_midpoint, 
                      ratio = y1bins$bin_value/y2bins$bin_value)

ggplot(plotdat, aes(x = x, y = ratio)) +
  geom_point() +
  scale_x_log10(limits = c(0.5, 250)) +
  scale_y_log10(limits = c(0.5, 250)) +
  geom_abline(slope = 1, intercept = log10(1/2), color = 'red') +
  theme_bw() +
  theme(aspect.ratio = 1)

```

Here is the ratio plotted directly, along with the actual power laws.

```{r}
f <- function(x, alpha1, alpha2, xmin1, xmin2) dpareto1(x,alpha1,xmin1)/dpareto1(x,alpha2,xmin2)

ggplot(data.frame(x=c(1,100))) +
  stat_function(fun=dpareto1, args=list(shape=1, min=1), color = 'red') +
  stat_function(fun=dpareto1, args=list(shape=2, min=1), color = 'blue') +
  stat_function(fun=f, args=list(alpha1=1,alpha2=2,xmin1=1,xmin2=1)) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  scale_x_log10() +
  scale_y_log10()

```

\newpage

# Ratio of two log-normals: Derivation

The ratio of two log-normal distributions has a log slope that increases as $O(\log n)$, as opposed to the constant log slope, or $O(1)$, shown above for the ratio of two power laws. Therefore we can say that the ratio of two log-normal distributions increases faster than a power law.

The derivation is as follows. First this is the definition of the log-normal probability density function. Note in all cases $\log$ is the natural logarithm.

$$P = \frac {e^{\frac{-(\log x - \mu)^2}{2\sigma^2}}}{x\sigma\sqrt{2\pi}}$$

The function we are interested in is the ratio of two of these distributions, each with its own mean and standard deviation.

$$y = \frac{P_1}{P_2} = \frac{\frac {e^{\frac{-(\log x - \mu_1)^2}{2\sigma_{1}^2}}}{x\sigma_1\sqrt{2\pi}}}{\frac {e^{\frac{-(\log x - \mu_2)^2}{2\sigma_2^2}}}{x\sigma_{2}\sqrt{2\pi}}}$$

Simplify.

$$y = \frac{\sigma_2}{\sigma_1}e^{\frac{(\log x - \mu_2)^2}{2\sigma_2^2}-\frac{(\log x - \mu_1)^2}{2\sigma_1^2}}$$

The slope of this function in log space is defined as:

$$ \frac {d \log y}{d \log x} = \frac{x}{y}\frac{dy}{dx}$$

Find $\frac{dy}{dx}$ with the chain rule. Because of the exponential function, it contains $y$ which makes our life easier since that will cancel out later.

$$\frac{dy}{dx} = y\Big(\frac{\log x - \mu_2}{\sigma_2^2x}-\frac{\log x - \mu_1}{\sigma_1^2x}\Big)$$

Next find $\frac{x}{y}\frac{dy}{dx}$.

$$\frac{x}{y}\frac{dy}{dx} = \frac{xy}{y}\Big(\frac{\log x - \mu_2}{\sigma_2^2x}-\frac{\log x - \mu_1}{\sigma_1^2x}\Big) = \frac{\log x - \mu_2}{\sigma_2^2}-\frac{\log x - \mu_1}{\sigma_1^2}$$

Combine like terms.

$$\frac{x}{y}\frac{dy}{dx} = \Big(\frac{\sigma_1^2 - \sigma_2^2}{\sigma_1^2\sigma_2^2}\Big)\log x + \Big(\frac{\sigma_2^2\mu_1 - \sigma_1^2\mu_2}{\sigma_1^2\sigma_2^2}\Big)$$

As you can see this is a constant times $\log x$ plus another constant. Since $\sigma_1 > 0$ and $\sigma_2 > 0$, the first constant is positive if $\sigma_1 > \sigma_2$. Therefore as $x$ gets large the slope increases as $O(\log x)$.
