---
title: "Replacing slow with fast"
author: "Quentin D. Read"
date: "July 16, 2019"
output: pdf_document
---

First load the data.
```{r load_data, message = FALSE, warning = FALSE}
library(tidyverse)

gdrive_path <- '~/google_drive/ForestLight'

load(file.path(gdrive_path, 'data/rawdataobj_alternativecluster.r'))

dat1995 <- alltreedat[[3]] %>%
  select(sp, fg, wsg, height_corr, dbh_corr, agb_corr) %>%
  mutate(fg = if_else(is.na(fg), 'unclassified', paste0('fg', fg)))

fgnames <- c('fast', 'pioneer', 'slow', 'breeder', 'medium', 'unclassified')
```


This is the allometric equation we used to get aboveground biomass from diameter, height, and wood specific gravity (WSG) of an individual tree.

```{r allometry}
agb.allometry <- function(wsg, dbh, height) { 0.0509 * wsg * ((dbh)^2) * height }
```


The only species-level trait that is used in the aboveground biomass allometry is WSG. This trait differs quite a bit among the functional groups.

```{r plot_wsg}
dat1995 %>%
  group_by(sp) %>%
  slice(1) %>%
  ungroup %>%
  ggplot(aes(x = factor(fg), y = wsg)) +
    geom_boxplot(fill = 'oldlace') + 
    geom_jitter() +
    scale_x_discrete(labels = fgnames) +
    labs(x = 'Functional group', y = 'Wood specific gravity') +
    theme_bw()
```

Here are the median (weighted by individual) WSG values for the functional groups.

```{r median_wsg}
(wsg_by_fg <- dat1995 %>%
  group_by(fg) %>%
  summarize(median_wsg = median(wsg)) )
```

Calculate an alternative allometry where all the slow individuals are replaced with a median fast individual of the same size. Sum up the "true" and "alternative" biomass by functional group.

```{r alternative}
fast_wsg <- wsg_by_fg$median_wsg[wsg_by_fg$fg == 'fg1']

dat1995 <- dat1995 %>%
  mutate(wsg_alternative = if_else(fg == 'fg3', fast_wsg, wsg),
         agb_alternative = agb.allometry(wsg_alternative, dbh_corr, height_corr))

biomass_sums <- dat1995 %>%
  group_by(fg) %>%
  summarize(biomass_true = sum(agb_corr),
            biomass_alternative = sum(agb_alternative)) 

# Rename FG3 to FG1 in alternative scenario
biomass_sums$biomass_alternative[1] <- biomass_sums$biomass_alternative[1] + biomass_sums$biomass_alternative[3]
biomass_sums$biomass_alternative[3] <- 0
```

Here is the result. It is notable that so much biomass is taken up by the long-lived pioneers even though they are relatively few individuals compared to the slow trees. There are just a lot of really big ones. This may be real or possibly partially due to the allometry overestimating biomass of big individuals.

```{r plot_biomass}
biomass_sums %>%
  mutate(fg = factor(fgnames, levels = fgnames)) %>%
  gather(scenario, biomass, -fg) %>%
  mutate(scenario = factor(scenario, levels = c('biomass_true','biomass_alternative'), labels = c('true','alternative'))) %>%
  ggplot(aes(x = scenario, y = biomass, fill = fg)) +
    geom_col(position = 'stack') +
    theme_bw() +
    scale_fill_manual(values = c(RColorBrewer::brewer.pal(5, 'Set1'), 'black')) +
    labs(y = 'Total aboveground biomass (kg)')
```

Overall, the aboveground biomass of the BCI plot would go down by 11% if all slow individuals were replaced with fast.

```{r calc_decrease}
1 - sum(biomass_sums$biomass_alternative) / sum(biomass_sums$biomass_true)
```