---
title: "Fit to central segment of total production"
author: "Quentin D. Read"
date: "April 3, 2018"
output: pdf_document
---

Here is a fit to the total production bins. It is not the recommended way of fitting a power law because it is a log-log linear regression fit to the binned values. However I don't know whether it's possible to fit it as a true power law since the definition of total production does involve creating size classes somehow. I would not be super confident in including these methods in the paper. Honestly I think it would be okay to just state qualitatively that the region in the middle has a slight positive slope (not really approximate energy equivalence as it is clearly a positive upward trend).

I fit the linear model through the binned values for (1) all 20 bins and (2) the bins that I eyeballed as being in the central flatter portion (about 4 cm to 40 cm dbh). Here, all trees in 1995 are used.

## Load data

```{r, warning = FALSE, message = FALSE}
fp <- 'C:/Users/Q/google_drive/ForestLight/data/data_20mar2018' ## CHANGE PATH AS NEEDED

# Read all the csvs in directory.
for (i in dir(fp, pattern = '.csv')) {
  n <- gsub('.csv','',i)
  assign(n, read.csv(file.path(fp, i), stringsAsFactors = FALSE))
}

library(dplyr)
library(rstan)
library(purrr)
library(ggplot2)
options(mc.cores = 3)

dat <- obs_totalprod %>%
  filter(year == 1995, fg == 'all')
```

## Define model

The model is defined in Stan so that we can get the Bayesian 95% credible intervals around the parameter estimates and the 95% predictive intervals around the fits.

```{r}
linear_model <- '
data {
  int<lower=0> N;
  vector<lower=0>[N] x;
  vector<lower=0>[N] y;
}
parameters {
  real beta0;
  real beta1;
  real<lower=0> sigma;
}
model {
  beta0 ~ normal(0, 10);
  beta1 ~ normal(0, 10);
  {
    vector[N] mu;
    for (i in 1:N) mu[i] = beta0 + beta1 * x[i];
    y ~ normal(mu, sigma);
  }
}
generated quantities {
  vector[N] ypred;
  for (i in 1:N) ypred[i] = (10^beta0) * (10^x[i])^beta1;
}
'
```

## Fit model to all 20 bins

```{r, warning = FALSE, message = FALSE, results = 'hide'}
dat_linear <- with(dat, list(N = nrow(dat),
                             x = log10(bin_midpoint),
                             y = log10(bin_value)))

mod_linear <- stan_model(model_code = linear_model)

fit_linear <- sampling(mod_linear, 
                       data = dat_linear, 
                       chains = 3, 
                       iter = 6000, 
                       warmup = 5000, 
                       seed = 30303)
sum_linear <- summary(fit_linear)
```

## Fit model to only the central segment

There are 9 bins in the central segment, based on eyeballing (bins 5 through 13 or roughly 4 cm dbh to 40 cm dbh).

```{r, warning = FALSE, message = FALSE, results = 'hide'}
dat_linear_straightportion <- with(dat[5:13,], list(N = length(5:13),
                             x = log10(bin_midpoint),
                             y = log10(bin_value)))
fit_linear_straight <- sampling(mod_linear, 
                                data = dat_linear_straightportion, 
                                chains = 3, 
                                iter = 6000, 
                                warmup = 5000, 
                                seed = 30303)
sum_linear_straight <- summary(fit_linear_straight)
```

## Credible intervals around parameters

The median slope estimate for fitting to all the data is -0.385, with 95% credible interval [-0.757, -0.004].  
The median slope estimate for fitting the central portion is +0.248, with 95% credible interval [+0.163, +0.338].

```{r}
sum_linear[[1]][c('beta0', 'beta1'), ]
sum_linear_straight[[1]][c('beta0', 'beta1'), ]
```

## Plot

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.height = 5, fig.width = 5}
area_core <- 42.84

# Extract quantiles of fitted values
pred1 <- extract(fit_linear)$ypred
pred2 <- extract(fit_linear_straight)$ypred

getpred <- function(x) {
x %>%
  as.data.frame %>%
  map(~ quantile(., probs = c(0.025, 0.5, 0.975))) %>%
  do.call(rbind, .) %>%
  as.data.frame %>%
  setNames(nm = c('q025', 'q50', 'q975'))
}

qpred1 <- data.frame(fit_to = 'all', dbh = dat$bin_midpoint, getpred(pred1))
qpred2 <- data.frame(fit_to = 'central section', dbh = dat$bin_midpoint[5:13], getpred(pred2))
qpred <- rbind(qpred1, qpred2)

ggplot(dat, aes(x = bin_midpoint)) + 
  geom_ribbon(data = qpred, aes(x = dbh, ymin = q025/area_core, ymax=q975/area_core, fill = fit_to), alpha = 0.2) +
  geom_line(data = qpred, aes(x = dbh, y = q50/area_core, color = fit_to)) +
  geom_point(aes(y = bin_value/area_core)) +
  theme_classic() +
  theme(legend.position = 'bottom') +
  scale_x_log10(name = 'Diameter') +
  scale_y_log10(name = 'Biomass/year/hectare')
  
```

