---
title: "Some forest light bar plots"
author: "Quentin D. Read"
date: "January 24, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setting up to make plots

First, in code not shown here, I load all the data and combine it with the different shade tolerance categories.

```{r, message = FALSE, warning = FALSE}
source('~/GitHub/forestlight/code/mergeshade.r')
```

Set the number of bins for the biomass classification.

```{r}
n_bins <- 10
```

Combine the biomass estimates from diameter at breast height with the estimates from diameter at ground height. Calculate the biomass increment between years 2 and 3 to use as the rate of biomass gain (production). Use biomass in year 3 to bin the trees. Split the data up by site so that it can be binned separately for each site.

```{r}
pdat <- dat %>% 
  mutate(biomass = pmax(BiomassDGH_year3_kg, BiomassDBH_year3_kg, na.rm=TRUE),
         biomass2 = pmax(BiomassDGH_year2_kg, BiomassDBH_year2_kg, 0, na.rm=TRUE),
         massprod = pmax(biomass - biomass2, 0, na.rm=TRUE)) %>% 
  select(Site, biomass, massprod, CII)
bcidat <- pdat %>% filter(Site == 'Barro_Colorado_Island', !is.na(CII), !is.na(biomass))
harvdat <- pdat %>% filter(Site == 'Harvard_Forest_LTER', !is.na(CII), !is.na(biomass))
```

Bin the trees at each site into bins based on the biomass. This means the scale of biomass is cut into segments of roughly equal length. The smaller size classes will have more individuals in them than the larger ones. The arithmetic scale is used rather than the logarithmic.

```{r}
bci_bin <- with(bcidat, cut(biomass, breaks = n_bins))
harv_bin <- with(harvdat, cut(biomass, breaks = n_bins))

bcidat$biomass_bin <- bci_bin
harvdat$biomass_bin <- harv_bin

```

This is a little code to make the labels of the bins more legible.

```{r}
# Pull numbers from bin labels
binlab2n <- function(x, islog) {
  lx <- levels(x)
  lxs <- strsplit(lx, ',')
  mins <- sapply(lxs, '[', 1)
  maxs <- sapply(lxs, '[', 2)
  mins <- as.numeric(substr(mins, 2, nchar(mins)))
  maxs <- as.numeric(substr(maxs, 1, nchar(maxs) - 1))
  if (islog) {
    mins <- round(10^mins,1)
    maxs <- round(10^maxs,1)
  }
  c(paste('<', maxs[1]), 
    paste(mins[2:(length(mins)-1)], 'to', maxs[2:(length(maxs)-1)]), 
    paste('>', mins[length(mins)]))
}
```

Next, set axes and plotting theme that can be applied to multiple plots.

```{r}
library(ggplot2)

sc_y <- scale_y_continuous(limits = c(0,5.5), expand = c(0,0))
sc_y2 <- scale_y_continuous(name = 'Biomass production (kg/y)', 
                            limits = c(0,3500), expand = c(0,0))
sc_x <- scale_x_discrete(name = 'Biomass (kg)', drop = FALSE)
th_bar <- theme_bw() + 
  theme(panel.grid = element_blank(), axis.text.x = element_text(size=6))
```

\newpage

# Plots

## CII plots

The plots of CII are "wrong" because they average the CII in each bin. This means an arbitrary number is treated as a real continuous measure when it's really an ordinal category. Obviously this is not good for modeling purposes but fine for visualization purposes. They look as expected.

```{r}
levels(bcidat$biomass_bin) <- with(bcidat, binlab2n(biomass_bin, islog=F))
levels(harvdat$biomass_bin) <- with(harvdat, binlab2n(biomass_bin, islog=F))

ggplot(bcidat, aes(x = biomass_bin, y = CII)) +
  stat_summary(geom = 'bar', fun.y = 'mean') +
  stat_summary( geom = 'errorbar', fun.data = 'mean_se', width = 0) +
  sc_y + sc_x + th_bar + ggtitle('BCI')

ggplot(harvdat, aes(x = biomass_bin, y = CII)) +
  stat_summary(geom = 'bar', fun.y = 'mean') +
  stat_summary( geom = 'errorbar', fun.data = 'mean_se', width = 0) +
  sc_y + sc_x + th_bar + ggtitle('Harvard Forest')
```

\newpage

## Production plots

Here are the biomass production plots, created by summing the biomass production in each bin. I did not show error bars since there is only a single value per bin. Not exactly energy equivalence, I would say!

```{r}
ggplot(bcidat, aes(x = biomass_bin, y = massprod)) +
  stat_summary(geom = 'bar', fun.y = 'sum') +
  sc_y2 + sc_x + th_bar + ggtitle('BCI')

ggplot(harvdat, aes(x = biomass_bin, y = massprod)) +
  stat_summary(geom = 'bar', fun.y = 'sum') +
  sc_y2 + sc_x + th_bar + ggtitle('Harvard Forest')
```

\newpage

## Data underlying the plots

For reference, here is the data used to generate each plot, along with numbers of trees in each bin. 

```{r}
# BCI
bcidat %>% 
  group_by(biomass_bin) %>% 
  summarize(mean_cii=mean(CII), total_production=sum(massprod), n=n())

# Harvard
harvdat %>% 
  group_by(biomass_bin) %>% 
  summarize(mean_cii=mean(CII), total_production=sum(massprod), n=n())

```

